Software Citation Principles

### 2016-07-23 sc-principles/rebuttal/rebuttal.tex

\niemnote{The idea of citations is not going to go away---but format and how they fit in ecosystem may change.}

\niemnote{Arfon: can you help here?}

\item \textbf{Reviewer comment:}
\emph{line 238: ``Citation is a record of software that is important to a research outcome'': This is another important issue, but not much more is said about this. What makes a software ``important'' to a research outcome? Or, more generally, what is the semantics of a software citation?}

\textbf{Our response:}
We have revised this discussion to better address these issues. In particular, we
added some references that discuss the importance of specific software versions to
research results (including data). For a particular study, it is up the authors to
determine which software is important.
\niemnote{how is this, including the new text?}

### 2016-07-23 sc-principles/software-citation-principles.tex
rearranged some text in 6.1
some edits at introduction

\niemnote{Do we want a better introduction, to help with the structure of the paper?}
\katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}

\niemnote{How is this?}
The main contribution of this document are the software citation principles,
written fairly concisely in this section and discussed further later in the
document (\S\ref{sec:discussion}).

The principles in this section are written fairly concisely, and discussed
further later in this document (\S\ref{sec:discussion}). Here, for example,
we do not define what software should be cited, but how it should be
cited, and we talk about how such decisions might be made in the
discussion section~(\S\ref{sec:discussion}).

@@

Some software which is, or could be, captured as part of data provenance may not be cited.
Citation is partly a record of software important to a research outcome\footnote{Citation
can be used for many purposes, including for software: which software has been used in
the work, which software has influenced the work, which software is the work superseding,
which software is the work disproving, etc.}, where provenance is a record of all steps
(including software) used to generated particular data within the research process.
This implies that for a data research product, provenance data will include some of
the cited software.
Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
and thus may depend on the specific version
used~\cite{Sandve:2013gh,Wilson:2014aa}---furthermore, errors in software or
environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
Similarly, the software metadata that is recorded as part of data provenance should be
a superset of the metadata recorded as part of software citation for the software that
was used in the work. The data recorded for reproducibility should also be a superset
of the metadata recorded as part of software citation.
These statements may also be true for software products.
In general, we intend the software citation principles to cover the minimum of what is
necessary for software citation for the purpose of software identification.
Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }

>>>>

\niemnote{Do we want a better introduction, to help with the structure of the paper?}
\katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}
\niemnote{How is this?}

The main contribution of this document are the software citation principles,
written fairly concisely in this section and discussed further later in the
document (\S\ref{sec:discussion}). In addition, we also motivate the creation
of these principles (\S\ref{sec:intro}), describe the process by which they
were created (\S\ref{sec:process}), summarize use cases related to software
citation (\S\ref{sec:use_cases}), and review related work
(\S\ref{sec:related_work}). We also lay out the necessary future work
(\S\ref{sec:futurework}).
%The principles in this section are written fairly concisely, and discussed
%further later in this document (\S\ref{sec:discussion}). Here, for example,
%we do not define what software should be cited, but how it should be
%cited, and we talk about how such decisions might be made in the
%discussion section~(\S\ref{sec:discussion}).

@@

Some software which is, or could be, captured as part of data provenance may not be cited.
Citation is partly a record of software important to a research outcome\footnote{Citation
can be used for many purposes, including for software: which software has been used in
the work, which software has influenced the work, which software is the work superseding,
which software is the work disproving, etc.}, where provenance is a record of all steps
(including software) used to generated particular data within the research process.
Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
and thus may depend on the specific version
used~\cite{Sandve:2013gh,Wilson:2014aa}---furthermore, errors in software or
environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
This implies that for a data research product, provenance data will include some of
the cited software.
Similarly, the software metadata that is recorded as part of data provenance should be
a superset of the metadata recorded as part of software citation for the software that
was used in the work. The data recorded for reproducibility should also be a superset
of the metadata recorded as part of software citation.
These statements may also be true for software products.
In general, we intend the software citation principles to cover the minimum of what is
necessary for software citation for the purpose of software identification.
Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
